---
title: "Project_3"
author: "Emerald Orcas"
date: "3/17/2020"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(RCurl)
library(stringr)
knitr::opts_chunk$set(echo = TRUE)
```

## Project 3 : Valuable Data Scientist Skills

The purpose of Project 3 is to use data to answer the question, "Which are the most valued data science skills?". 

After initial brainstorming, we determined to use the following dataset to perform analysis: https://www.kaggle.com/elroyggj/indeed-dataset-data-scientistanalystengineer/kernels.

The dataset takes data science job postings and breaks them down by industry, company, skills, job titles, location and categorical job types (data scientist, data analyst, and data engineer). 

We determind this is an appropriate dataset since it is employers that generally determine what are valued skill sets in an occupation.


```{r}
#Import indeed dataset

url <- getURL("https://raw.githubusercontent.com/geeman1209/MSDATA2020/master/DATA607/Project_3/indeed_job_dataset.csv")

raw_data <- read.csv(text = url)


#get rid of unnecessary columns
df_1 <- raw_data
drops <- c("X","Link","Company","No_of_Reviews","No_of_Stars","Date_Since_Posted","Company_Revenue","Company_Employees", "Description", "Location")
df_1 <- select(df_1 ,-c(drops))

#delete the columns 17-27

df_2 <- select(df_1, -c(17:27))


#add single column for Junior or Senior Data Scientist in Job Title
df_2['Level'] <- ifelse(grepl("Jr|Junior",ignore.case=TRUE,df_2$Job_Title),'Junior',ifelse(grepl("Sr|Senior",ignore.case=TRUE,df_2$Job_Title),"Senior","Other"))

#add single column for industry
df_2['Industry'] <- ifelse(df_2$Consulting.and.Business.Services==1,'Consulting&Business',ifelse(df_2$Internet.and.Software,'Internet&Software',ifelse(df_2$Banks.and.Financial.Services==1,'Banks&FiancialServices',ifelse(df_2$Health.Care==1,'HealthCare',ifelse(df_2$Insurance==1,'Insurance',ifelse(df_2$Other_industries==1,'Other',NA))))))

```

# Data Analysis  
What are the most popular skills for a data scientist? Do these top skils differ for data analysts and data engineers?  
```{r most-common-skills}
skills <- df_2 %>% group_by(Job_Type) %>% summarise(sum(python),sum(sql),sum(machine.learning),sum(r),sum(hadoop),sum(tableau),sum(sas),sum(spark),sum(java),sum(Others))

colnames(skills) <- c('Job_Type','Python','SQL','Machine_Learning','R','Hadoop','Tableau','SAS','Spark','Java','Others')

skills <- pivot_longer(skills,cols=c(2:11),names_to = "Skill",values_to = "Count")

ggplot(skills, aes(x=reorder(Skill,-Count),y=Count,fill=Job_Type)) +
  geom_bar(stat="identity", position = 'dodge') +
  coord_flip() +
  xlab('Skill') +
  ylab('Count of Postings that Listing this Skill') + 
  ggtitle('Top 10 Skills Among Data Job Listings - LinkedIn') +
  theme(plot.title = element_text(hjust = 0.5))
```
Do Skills Differ Among Junior vs. Senior Data Scientists?
```{r data-science-skills}
ds <- filter(df_2,Job_Type == 'data_scientist',Level %in% c("Senior","Junior"))
ds_skills <- ds %>% group_by(Level) %>% summarise(sum(python),sum(sql),sum(machine.learning),sum(r),sum(hadoop),sum(tableau),sum(sas),sum(spark),sum(java),sum(Others))

colnames(ds_skills) <- c('Level','Python','SQL','Machine_Learning','R','Hadoop','Tableau','SAS','Spark','Java','Others')

ds_skills_long <- pivot_longer(ds_skills,cols=c(2:11),names_to = "Skill",values_to = "Count")

ggplot(ds_skills_long, aes(x=reorder(Skill,Count),y=Count,fill=Level)) +
  geom_bar(stat="identity", position = 'dodge') +
  coord_flip() +
  xlab('Skill') +
  ylab('Count of Postings that Listing this Skill') + 
  ggtitle('Top 10 Skills Among Data Job Listings - LinkedIn') +
  theme(plot.title = element_text(hjust = 0.5))
```
Which industries have the most job postings for data scientists?
```{r count-job-posting-by-industry}
posts_by_industry <- filter(df_2,Job_Type == 'data_scientist',is.na(df_2$Industry) == FALSE)
industry_count <- posts_by_industry %>% group_by(Industry) %>% tally()

colnames(industry_count) <- c('Industry','Count')

ggplot(industry_count, aes(x=reorder(Industry,Count),y=Count)) +
  geom_bar(stat="identity", position = 'dodge') +
  coord_flip() +
  xlab('Industry') +
  ylab('Count of Job Postings') + 
  ggtitle('Top 10 Skills Among Data Job Listings - LinkedIn') +
  theme(plot.title = element_text(hjust = 0.5))
```

Do certain industries require certain skills? We're going to ignore OTHER since it's overwhelmingly popular and we don't know which industries it represents.
```{r industry-skills}
#remove rows where industry is NA or OTHER
industry <- filter(df_2,Job_Type == 'data_scientist',df_2$Industry != 'Other',is.na(df_2$Industry) == FALSE)
industry_skills <- industry %>% group_by(Industry) %>% summarise(sum(python),sum(sql),sum(machine.learning),sum(r),sum(hadoop),sum(tableau),sum(sas),sum(spark),sum(java),sum(Others))

colnames(industry_skills) <- c('Industry','Python','SQL','Machine_Learning','R','Hadoop','Tableau','SAS','Spark','Java','Others')

industry_skills <- pivot_longer(industry_skills,cols=c(2:11),names_to = "Skill",values_to = "Count")

#Internet & Software
internet_consulting <- filter(industry_skills,industry_skills$Industry %in% c('Internet&Software','Consulting&Business'))
a <- ggplot(internet_consulting, aes(x=reorder(Skill,-Count),y=Count,fill=Industry)) +
  geom_bar(stat="identity", position = 'dodge')+
  xlab('') +
  ylab('Count of Job Posts') +
  theme(axis.text.x = element_text(angle = 40))

#Healthcare, Insurance & Banks
healthcare_insurance_banks <- filter(industry_skills,industry_skills$Industry %in% c('HealthCare','Insurance','Banks&FinancialServices'))
b <- ggplot(healthcare_insurance_banks, aes(x=reorder(Skill,-Count),y=Count,fill=Industry)) +
  geom_bar(stat="identity", position = 'dodge')+
  xlab('Skill') +
  ylab('Count of Job Posts') +
  theme(axis.text.x = element_text(angle = 40))

grid.arrange(a,b,nrow=2,top = "Top Skills Across Industry for Data Scientist Job Postings")

```
# Which skills pay the highest salary?
```{r avg-highest-salary}
ds_salary <- select(filter(df_2,df_2$Job_Type=='data_scientist'), Queried_Salary,python,sql,machine.learning,r,hadoop,tableau,sas,spark,java,Others)
ds_salary_job_type <- pivot_longer(ds_salary,cols=c(2:11),names_to='Skill')
skill <- ds_salary_job_type %>% group_by(Queried_Salary,Skill) %>% summarise(sum(value))

ggplot(skill, aes(x=Queried_Salary,y='sum(value)',fill=Skill)) +
  geom_bar(stat="identity", position = 'dodge')+
  xlab('Skill') +
  ylab('Count of Job Posts') +
  theme(axis.text.x = element_text(angle = 40))
```

# Including Plots


```{r pressure, echo=TRUE}
plot(pressure)
```
