---
title: "Project_3"
author: "Emerald Orcas"
date: "3/17/2020"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(RCurl)
knitr::opts_chunk$set(echo = TRUE)
```

## Project 3 : Valuable Data Scientist Skills

The purpose of Project 3 is to use data to answer the question, "Which are the most valued data science skills?". 

After initial brainstorming, we determined to use the following dataset to perform analysis: https://www.kaggle.com/elroyggj/indeed-dataset-data-scientistanalystengineer/kernels.

The dataset takes data science job postings and breaks them down by industry, company, skills, job titles, location and categorical job types (data scientist, data analyst, and data engineer). 

We determind this is an appropriate dataset since it is employers that generally determine what are valued skill sets in an occupation.


```{r}
#Import indeed dataset

url <- getURL("https://raw.githubusercontent.com/geeman1209/MSDATA2020/master/DATA607/Project_3/indeed_job_dataset.csv")

raw_data <- read.csv(text = url)


#get rid of unnecessary columns
df_1 <- raw_data
drops <- c("X","Link","Company","No_of_Reviews","No_of_Stars","Date_Since_Posted","Company_Revenue","Company_Employees", "Description", "Location")
df_1 <- select(df_1 ,-c(drops))

#delete the columns 17-27

df_2 <- select(df_1, -c(17:27))


```

## Including Plots


```{r pressure, echo=TRUE}
plot(pressure)
```
