---
title: "Data 605 - WK4"
author: "Gabe Abreu"
date: "9/20/2020"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data 605 - Assignment 4

###Problem Set 1

Write code to in R to compute X = A(A^T) and Y = (A^T)A
```{r code1}
#Create the matrix A
A <- matrix(c(1,2,3,-1,0,4), byrow = T, nrow = 2)
A


#Compute X by multiplying A by its transpose
X <- A%*%t(A)
X

#Compute Y by multiplying the transpose by A
Y <- t(A)%*%A
Y
```
Compute the eigenvalues and eigenvectors of X and Y using the built-in commands in R.
```{r code2}
eigX <- eigen(X)
eigY <- eigen(Y)
```

Eigvenvalues:
```{r 1}
#Eigenvalues for X
eigX$values

#Eigenvalues for Y
eigY$values
```

Eigenvectors:
```{r 2}
#Eigen vectors for x
eigX$vectors


#Eigenvectors for Y
eigY$vectors
```

Compute the the left-singular/singular values/right-singular vectors using svd command
```{r 3}
SVD_A <- svd(A)

sing_val <- SVD_A$d #singular values of the given matrix

left_val <- SVD_A$u #left singular vectors of matrix

right_val <- SVD_A$v #right singular vectors of matrix
```
Examine the two sets of singular vectors and show that they are indeed eigenvectors of X and  Y.
```{r 4 }
#Compare Eigenvectors of X and left values of A
list(eigX$vectors, left_val)


#Compare Eigenvectors of Y and right values of A
list(eigY$vectors, right_val)
```

In addition, the two non-zero eigenvalues (the 3rd value will be very close to zero, if not zero) of both X and Y are the same and are squares of the non-zero singular values of A.
```{r 5}
#Let's look at the single values without them being squared
sing_val

#Single values squared
sq_values <- round(sing_val^2)
sq_values

round(eigX$values)

round(eigY$values)
```

## Problem Set 2

```{r code}

myinverse <- function(A) {
    #Check if matrix is square
    if(dim(A)[1] != dim(A)[2]){
        return("Not a square matrix")
    }
    
    #If the matrix is 2x2, this will return the inverse
    if(dim(A)[1] == 2)
    {
       det_1 <- (1/((A[1,1]*A[2,2])-(A[1,2]*A[2,1])))
       
       A_2 <- matrix(c(A[2,2], -A[1,2], -A[2,1], A[1,1]), byrow = T, nrow = 2)
       
       return (det_1 * A_2)
    }
    
    
    #Create Co-factor Matrix 
    C_Matrix <- matrix(rep(0,length(A)),ncol=ncol(A))
    
    
    for (i in 1:ncol(A)) {
        for (j in 1:nrow(A)) {
            C_Matrix[i,j] <- (-1)^(i+j)*det(A[-i,-j])
        }
    }
    
    #Inverse of A is equal to transpose of Co-Factor Matrix divided by the determinant of A
    B <- t(C_Matrix)/det(A)
    
    return(B)
    
}

```


Let's run some test scenarios
```{r code2}
#Test matrix was taken from https://www.mathsisfun.com/algebra/matrix-inverse.html
test_matrix <- matrix(c(3, 3.5, 3.2, 3.6), byrow = T, nrow = 2)
test_matrix

test_matrix2 <- matrix(c(6,1,1,4,-2,5,2,8,7), byrow = T, nrow = 3)
test_matrix2

#Use the built-in solve function (also the website verifies the answer)
solve(test_matrix)
solve(test_matrix2)


#Test function
myinverse(test_matrix)
myinverse(test_matrix2)
```
