---
title: "Data605_Assignment7_GA"
author: "Gabe Abreu"
date: "10/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data 605 Assignment # 7 

1. Let X1, X2, . . . , Xn be n mutually independent random variables, each of
which is uniformly distributed on the integers from 1 to k. Let Y denote the
minimum of the Xi’s. Find the distribution of Y .

Answer:

Since Y is equivalent to the minimum of the Xi's, there is an "k" amount of possibilities that Xi can take: 1, 2, 3,...k. The total possible number of assignments for the collection of random variables  is k^(n). 


When Y = 1, the way to get 1 is k^(n) - (k - 1)^n, with k^n representating the total possibilities and (k-1)^n are the options where none of the random variables equal 1.

P(X=1) = (k^(n) - ((k-1)^n))/ (k^(n))

P(X=2) = ((k^(n) - (k - 2)^(n) - [k^n - (k - 1)^n])/ (k^(n)) -->  (k-1)^n - (k-2)^(n) / (k)^n

P(X = i) = ((k - i + 1)^n - (k-i)^n) / (k)^n

## Part 2

2. Your organization owns a copier (future lawyers, etc.) or MRI (future doctors).

This machine has a manufacturer’s expected lifetime of 10 years. This means
that we expect one failure every ten years. (Include the probability statements
and R Code for each part.).

a. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a geometric. (Hint: the probability is equivalent to not failing during the first 8 years..)

```{r}
fail_p <- 1/10

#probability of the machine not failing 
success_p <- 1 - fail_p

#use p_geom formula to model geometric
geometric <-  1 - pgeom(8-1, fail_p)
geometric

#Expected value
value_geomtric <- 1/fail_p
value_geomtric

#Standard Deviation
sd_geo <- sqrt(success_p/(fail_p)^2)

sd_geo
```

b. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as an exponential. 

```{r}
lambda <- 1/10

expoential_p <- pexp(8, lambda, lower.tail = FALSE)
expoential_p


#Expected Value
expected_exp_val <- 1/lambda
expected_exp_val

#STD Value

expo_sd <- sqrt(1/(lambda^2))
expo_sd
```

c. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8 years) 
```{r}
p <- .1
complement <- 1 - p
success <- 0
n <- 8

binomial_p <- dbinom(success, n, p)
binomial_p


#Expected value
binomial_exp_value <- n * p
binomial_exp_value

# STD Value
binomial_sd <- sqrt(n * p * complement)
binomial_sd
```
d. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a Poisson.
```{r}
lambda <- 8/10

ppois(0, lambda = lambda)

#Expected Value
Expected_pois_val <- lambda
Expected_pois_val

#Standard Deviation
standard_pois_dev <- sqrt(lambda)
standard_pois_dev
```
