---
title: "HW12"
author: "Gabe Abreu"
date: "11/15/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
```

## Data 605 #12


1. Provide a scatterplot of LifeExp~TotExp, and run simple linear regression. Do not transform the variables. Provide and interpret the F statistics, R^2, standard error,and p-values only. Discuss whether the assumptions of simple linear regression met.
```{r}
who_info <- read.csv('who.csv')

head(who_info)

simpleRegression <- lm(LifeExp ~ TotExp, data = who_info)
plot(who_info$TotExp, who_info$LifeExp)
abline(simpleRegression)
```


```{r}
summary(simpleRegression)



#Residuals
simple.res = resid(simpleRegression)

 plot(fitted(simpleRegression), simple.res) 
 abline(0,0)
 
 
hist(simple.res)

```

Q-Q Plots
```{r}
qqnorm(resid(simpleRegression))
qqline(resid(simpleRegression))

```

The F-Statistics suggests there is a relationship between the indepdent and dependent variables. The value of the F statistics is 65.26.

The R^2 represents how well the data fits the model which in this case the value of R-square (0.2577) is slow, indicating the data doesn't fit the model.

The residual standard error is the standard deviation of the residuals, the smaller the value the better. In in our model the residual standard error is high at 9.371

The p-value shows statistical significance at 7.71e-14.

As it stands, this is not a good model. 

2. Raise life expectancy to the 4.6 power (i.e., LifeExp^4.6). Raise total expenditures to the 0.06 power (nearly a log transform, TotExp^.06). Plot LifeExp^4.6 as a function of TotExp^.06, and r re-run the simple regression model using the transformed variables. Provide and interpret the F statistics, R^2, standard error, and p-values. Which model is "better?"
```{r}
who_info$LE2 <- who_info$LifeExp^4.6
who_info$TE2 <- who_info$TotExp^0.06

simpleRegression2 <- lm(who_info$LE2 ~ who_info$TE2, data = who_info)
plot(who_info$TE2, who_info$LE2)
abline(simpleRegression2)
```

```{r}
summary(simpleRegression2)

#Residuals
simple2.res = resid(simpleRegression2)

 plot(fitted(simpleRegression2), simple2.res) 
 abline(0,0)
 
 
hist(simple2.res)

```
Q-Q Plots
```{r}
qqnorm(resid(simpleRegression2))
qqline(resid(simpleRegression2))

```

The model is better based on the criteria used to judge linear models. 

The F statistics is much higher and p-value is still significant. The R^2 valueu is greatly improved. The residuals show normality and linearity. 

3. Using the results from 3, forecast life expectancy when TotExp^.06 =1.5. Then forecast life expectancy when TotExp^.06=2.5.

The formula is y = -736527910 + 620060216 * x
```{r}
predictor <- function(x) {
    y <- -736527910 + 620060216 * x
    y <- y^(1/4.6)
    
    return(y)
}

```

totExp^0.6 = 1.5
```{r}
predictor(1.5)
```
totExp^0.6 = 2.5
```{r}
predictor(2.5)
```

4. Build the following multiple regression model and interpret the F Statistics, R^2, standard error, and p-values. How good is the model?
LifeExp = b0+b1 x PropMd + b2 x TotExp +b3 x PropMD x TotExp

```{r}
multiRegression <- lm(LifeExp ~ PropMD + TotExp + PropMD*TotExp, data = who_info)
summary(multiRegression)

```

The model doesn't work as well as the second model. The residual standard error is high at 8.765. The R^2 value is low at 0.3574 and the F-Statistic is 34.49. The p-value is still significant being well below 0.05. 

5. Forecast LifeExp when PropMD=.03 and TotExp = 14. Does this forecast seem realistic? Why or why not?

```{r}
expdata <- data.frame(PropMD=0.03, TotExp=14)
round(predict(multiRegression, expdata))
```

The value is much too high at 107. Doesn't seem realistic given the rest of the data. 