---
title: "GA_HW11_Data605"
author: "Gabe Abreu"
date: "11/7/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(psych)
library(stats)
knitr::opts_chunk$set(echo = TRUE)
```

## Data 605 HW11

Using the “cars” dataset in R, build a linear model for stopping distance as a function of speed and replicate the analysis of your textbook chapter 3 (visualization, quality evaluation of the model, and residual analysis.)


```{r}
#Load the cars dataset into a variable as a data frame
cars <- datasets::cars


#Let's investigate the data
summary(cars)

describe(cars)


#Using the lm formula to build the linear regression model
cars.lm <- lm(dist ~ speed, data=cars)

#The summary will give us coefficients as well the slope and intercept that can be used to build the linear regression model evaluating the relationship between distance and speed.
summary(cars.lm)
```
What's important to take note is the p value which is less than 0.05, making it statistically significant. The intercept is -17.5791 and the slope is 3.9324, meaning the distance changes by 3.9324 per changes in speed.  

The adjusted R^2 is 0.6438. 

## Visualizing the Data
```{r}
#ggplot without getting the slope and intercept manually
ggplot(data = cars, aes(x = speed, y = dist)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE)

#ggplot with slope and intercept gotten from cars.lm

mx <- coef(cars.lm)[1]
slope <- coef(cars.lm)[2]

ggplot(cars.lm, aes(cars$speed, cars$dist)) +
  geom_point() +
  geom_abline(slope = slope, intercept = mx, show.legend = TRUE)
```

## Model Diagnostic

In order to evaluate whether the linear model is reliable, 3 things need to be accounted for
a)linearity
b)normal residuals
3)variability

Let's check for variability:
```{r}
ggplot(data = cars.lm, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("Fitted values") +
  ylab("Residuals")

```

There is no pattern in the plot above which is a sign of linearity. 

Check for Normal Residuals:
```{r}
ggplot(data = cars.lm, aes(x = .resid)) +
  geom_histogram(binwidth = 10) +
  xlab("Residuals")

qqnorm(resid(cars.lm))
qqline(resid(cars.lm))
```

The residuals seem close to normally distributed even though the histogram and Q-Q plot show outliers on the far right end. The residuals don't have to be perfectly normal but close enough. I would conclude that it is a viable model.