---
title: "Data605_Discussion605"
author: "Gabe Abreu"
date: "11/4/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

## Discussion 11

I'm going to use the built-in data set USArrests to demonstrate linear regression.

Is there a relationship between urban population and assaults?

```{r USArrests}
data('USArrests')
summary(USArrests)
```


The scatter plot can visualize linear relationships between the depdendent variable and indepdendent variable.
```{r}
head(USArrests)

scatter.smooth(x=USArrests$UrbanPop, y=USArrests$Assault, main="UrbanPop ~ Assault")
```

Let's build the linear regression and then print the summary. 



```{r}
linear_reg <- lm(UrbanPop ~ Assault, data = USArrests)
summary(linear_reg)

```
The p value is greater than 0.05, which means it's not statistically significant.

Plotting the linear regression with a scatter plot.
```{r}
ggplot(data = USArrests, aes(x = UrbanPop, y = Assault)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE)

```

Residual Plot
```{r}
assault.res = resid(linear_reg)

 plot(fitted(linear_reg), assault.res) 
 abline(0,0)
 
 
hist(assault.res)

```




Q-Q Plot

The plot below shows there are outliers at both ends of the dat, but more so towards upper quantiles. 
```{r}
qqnorm(resid(linear_reg))
qqline(resid(linear_reg))
```
