---
title: "HW3"
author: "Gabe Abreu"
date: "3/13/2021"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(plotly)
library(shiny)
library(rsconnect)
rsconnect::deployApp(appName='Module3Q1' )
knitr::opts_chunk$set(echo = TRUE)
```


## Question 1

As a researcher, you frequently compare mortality rates from particular causes across
different States. You need a visualization that will let you see (for 2010 only) the crude
mortality rate, across all States, from one cause (for example, Neoplasms, which are
effectively cancers). Create a visualization that allows you to rank States by crude mortality
for each cause of death.


```{r}

cdc_df <- read.csv('https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module3/data/cleaned-cdc-mortality-1999-2010-2.csv')

```

```{r}
ui <- fluidPage(
  headerPanel('States Ranked Per Crude Rate Per ICD'),
  sidebarPanel(
    selectInput('icd', 'ICD Chapter', unique(cdc_df$ICD.Chapter), selected='External causes of morbidity and mortality'),
    selectInput('Year', 'Year', unique(cdc_df$Year), selected=2010)
  ),
  mainPanel(
    plotOutput('plot1')
  )
)

server <- function(input, output) {
  
  output$plot1 <- renderPlot({
    
    dfSlice <- cdc_df %>%
      filter(ICD.Chapter == input$icd, Year == input$Year)
    
    ggplot(dfSlice, aes(x = Crude.Rate , y = reorder(State, Crude.Rate), fill= State)) +
      geom_bar(stat='identity') +
      ylab("State") + 
      theme(axis.text.y = element_text(size = 11, hjust = 1))
  })
  
  
}

shinyApp(ui = ui, server = server)

```