---
title: "GAbreu_Assign1"
author: "Gabe Abreu"
date: "8/30/2020"
output: html_document
---

```{r setup, include=FALSE}
library(pracma)
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 1 Data 605

1. Problem Set 1

You can think of vectors representing many dimensions of related information. For instance, Netflix might store all the ratings a user gives to movies in a vector. This is clearly a vector of very large dimensions (in the millions) and very sparse as the user might have rated only a few movies. Similarly, Amazon might store the items purchased by a user in a vector, with each slot or dimension representing a unique product and the value of the slot, the number of such items the user bought. One task that is frequently done in these settings is to find similarities between users. And, we can use dot-product between vectors to do just that. As you know, the dot-product is proportional to the length of two vectors and to the angle between them. In fact, the dot-product between two vectors, normalized by their lengths is called as the cosine distance and is frequently used in recommendation engines.

(1) Calculate the dot product u,v where u = [0.5;0.5] and v = [3:-4]

```{r}
u <- c(0.5, 0.5)
v <- c(3, -4)

dot_prod <- dot(u, v)
dot_prod
```

(2) What are the lengths of u and v?
```{r}
length_u <- sqrt((0.5^2) + (0.5^2))
length_u


length_v <- sqrt((3^2) + (-4 * -4))
length_v
```

(3) What is the linear combination: 3u- 2v?
```{r}

combo <- (3 * u) - (2 * v)
combo
    
```

(4) What is the angle between u and v?
```{r}
theta <- acos( dot_prod / (length_u * length_v))
theta

```

2. Problem Set 2

Set up a system of equations with 3 variables and 3 constraints and solve for x. Please write a function in R that will take two variables (matrix A & constraint vector b) and solve using elimination. Your function should produce the right answer for the system of equations for any 3-variable, 3-equation system. You don’t have to worry about degenerate cases and can safely assume that the function will only be tested with a system of equations that has a solution. Please note that you do have to worry about zero pivots, though. Please note that you should not use the built-in function solve to solve this system or use matrix inverses. The approach that you should employ is to construct an Upper Triangular Matrix and then back-substitute to get the solution. Alternatively, you can augment the matrix A with vector b and jointly apply the Gauss Jordan elimination procedure.

```{r}
#The method chosen is to contrcut an Upper Triangular Matrix and then utilize back-subsitution to get the solution.

elimination <- function(A, b){
    
    coeff_matrix <- cbind(A, b)
    
    #Start with row 1 of the co-efficient matrix 
    
    #Pivot: The first non-zero element in the row being evaluated
    pivot = coeff_matrix[1,1]
    
    #Multiplier: The element being eliminated divided by the Pivot
    multiplier = coeff_matrix[2,1]/pivot
    
    #Subtract Multiplier times row n from row n + 1//make 2,1 = 0
    r1 <- coeff_matrix[1,]
    r2 <- coeff_matrix[2,]
    
    coeff_matrix[2,] <- r2 - (r1 * multiplier)
    
    #Do it again for the third row// make 3,1 = 0
    pivot <- coeff_matrix[1,1]
    multiplier <- coeff_matrix[3,1]/pivot
    r3 <- coeff_matrix[3,]
    r1 <- coeff_matrix[1,]
    coeff_matrix[3,] <- r3 - (r1 * multiplier)
    
    
    #Find New pivot in second column/Second Row && make 3,2 = 0
    pivot <- coeff_matrix[2,2]
    multiplier <- coeff_matrix[3,2]/pivot
    r3 <- coeff_matrix[3,]
    r2 <- coeff_matrix[2,]
    coeff_matrix[3,] <- r3 - (r2 * multiplier)
    
    
    #The solution
    
    #Solve for the third variable
    x3 <- coeff_matrix[3,4]/coeff_matrix[3,3]
    
    #Substitute and solve for second variable
    x2 <- (coeff_matrix[2,4] - coeff_matrix[2,3] * x3) / coeff_matrix[2,2]
    
    #Substitute and solve for final variable in first row
    x1 <- (coeff_matrix[1,4] - coeff_matrix[1,3] * x3 - coeff_matrix[1,2] * x2) / coeff_matrix[1,1]
    
    return (rbind(x1, x2, x3))
    
}

test_case_a <- matrix(c(1,1,3,2,-1,5,-1,-2,4), byrow = T, 3, 3)
test_case_b <- c(1, 2, 6)

elimination(test_case_a, test_case_b)
```

References:
https://www.youtube.com/watch?v=jOC4fMgl7TI

https://stackoverflow.com/questions/54146330/angle-between-vector-and-list-of-vectors-in-r

An excellent video breaking down the gauss jordan method in python: https://www.youtube.com/watch?v=xOLJMKGNivU

