---
title: "Discussion5"
author: "Gabe Abreu"
date: "9/23/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 1.2

p.35 Exercise 1

Let Ω = {a, b, c} be a sample space. Let m(a) = 1/2, m(b) = 1/3, and
m(c) = 1/6. Find the probabilities for all eight subsets of Ω.


P(Ω) = 0 <- subset 8

```{r}
first_subset_a = 1/2
second_subset_b = 1/3
third_subset_c = 1/6

fourth_subset_ab = first_subset_a + second_subset_b
fourth_subset_ab


fifth_subset_ac = first_subset_a + third_subset_c
fifth_subset_ac

sixth_subset_bc = second_subset_b + third_subset_c
sixth_subset_bc


seventh_subset_abc = first_subset_a + second_subset_b + third_subset_c
seventh_subset_abc
```


p.35 Exercise 6

A die is loaded in such a way that the probability of each face turning up
is proportional to the number of dots on that face. (For example, a six is
three times as probable as a two.) What is the probability of getting an even
number in one throw?

```{r}

#The probability of getting 1 is p, 2 is 2p, 3, = 3p...6 = 6p
#Therefore, p + 2p + 3p + 4p + 5p + 6p = 21p
# Probability is 1/21 * face of the die

basic_die <- matrix(c(1,2,3,4,5,6), byrow = T, ncol = 3)
basic_die

#Die * probability 

probability <- basic_die * (1/21)
probability


even_probability <- probability[2,1] + probability[1,2] + probability[2,3]

paste(c("The probability of rolling an even number first is ", even_probability))
```


