---
title: 'Assignment #3'
author: "Gabe Abreu"
date: "2/15/2020"
output: html_document
---

```{r setup, include=FALSE}
library(glue)
library(stringr)
library(data.table)
library(plyr)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 3

### Question 1
Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r majors}

major_list <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"

Majors <- read.csv(major_list)

data_majors <- as_data_frame(Majors)

SelectedMajors <- data_majors$Major[grep("DATA|STATISTICS", data_majors$Major)]

View(SelectedMajors)

Majors %>% filter(str_detect(Major, ("DATA|STATISTICS")))

```
### Question 2
#2 Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry"

````{r}
fruits <- c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

dput(as.character(fruits))

`````

### Question 3
 Describe, in words, what these expressions will match:

(.)\1\1
Answer: same character appears 3 times in a row

"(.)(.)\\2\\1"
Answer: 2 characters attached to the same 2 characters in reverse order

(..)\1
Answer:Any 2 characters repeated

"(.).\\1.\\1"
Answer:1 character repeated three times with a different character in between every original character occurrence "rtrsr"

"(.)(.)(.).*\\3\\2\\1"
Answer:3 characters followed by zero or more characters followed by the original 3 characters in reverse order.

### Question 4
Construct regular expressions to match words that:

Start and end with the same character.
Regular Expression: "^(.)((.*\\1$)|\\1?$)"

Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
Regular Expression: "[A-Za-z][A-Za-z]).*\\1"

Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)
Regular Expression:"([a-z].*\\1.*\\1")

```{r test}
rando_words <- c("bell","apple","dog","eye","bob","test","mom","sense","church","banana","pepperoni","red","green","England","eleven","ten","twelve","soso","oso","bandana", "Louisiana", "Missouri", "Mississippi", "Connecticut", "google", "conscience")

#A
four_a <- str_subset(rando_words, "^(.)((.*\\1$)|\\1?$)")
four_a

#B
four_b <- str_subset(rando_words, "([A-Za-z][A-Za-z]).*\\1")
four_b
#C
four_c <- str_subset(rando_words, "([A-Za-z]).*\\1.*\\1")
four_c
```