---
title: "Discussion_12"
author: "Gabe Abreu"
date: "11/12/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(corrplot)
library(ggplot2)
library(stats)
library(Hmisc)
library(PerformanceAnalytics)
knitr::opts_chunk$set(echo = TRUE)
```

## Discussion 12

This data set was downloaded from Kaggle and sourced from the World Health Organization. I want to examine variables that are positively correlated with life expectancy. 

```{r}
LE_data <- read.csv('Life Expectancy Data.csv')

head(LE_data)

summary(LE_data)
```


```{r}
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}

LE_data_df <- LE_data[ ,4:length(LE_data)]

LE_data_df.cor = cor(LE_data_df, method = c("spearman"))

newCorr <- rcorr(as.matrix(LE_data_df))

flattenCorrMatrix(newCorr$r, newCorr$P)


```
We can see that life expectancy and schooling have a high correlation and a statistically significant p value.

## Including Plots


```{r}
corrplot(newCorr$r, type="upper", order="hclust", 
         p.mat = newCorr$P, sig.level = 0.05, insig = "blank")

chart.Correlation(LE_data_df, histograme=TRUE, pch=19)

palette = colorRampPalette(c("green", "white", "red")) (10)
heatmap(x = newCorr$r, scale="column", col = palette, symm = TRUE)

```

Schooling and life expectancy are positively correlated. Let's create a linear regression model and later evaluate it.

```{r}
lif <- LE_data_df$Life.expectancy
sch <- LE_data_df$Schooling

lif_sch_lm <- lm(sch ~ lif)


summary(lif_sch_lm)
```



```{r}
ggplot(data = LE_data_df, aes(x = Schooling, y = Life.expectancy)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE)

```
Examining the Residuals

```{r}
life_exp.res = resid(lif_sch_lm)

 plot(fitted(lif_sch_lm), life_exp.res) 
 abline(0,0)
 
 
hist(life_exp.res)



qqnorm(resid(lif_sch_lm))
qqline(resid(lif_sch_lm))


```