---
title: "GAbreu_Assignment2_PS2"
author: "Gabe Abreu"
date: "9/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data 602 Assignment 2 Part 2

### 2. Problem set 2

Matrix factorization is a very important problem. There are supercomputers built just
to do matrix factorizations. Every second you are on an airplane, matrices are being
factorized. Radars that track flights use a technique called Kalman filtering. At the heart
of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear
systems of equations when they track your flight using radars.
Write an R function to factorize a square matrix A into LU or LDU, whichever you
prefer.


```{r}
#Two test matrices
#The first matrix is taken from the week 2 lecture handout
A <- matrix(c(1,2,3,1,1,1,2,0,1), nrow=3)
B<- matrix(c(1,1,2,3), nrow = 2)
C <- matrix(c(1,2,3,4,-1,0,1,3,0,1,-2,1,5,4,-2,-3), nrow = 4)


LU_Decomp <- function(mat){

#Get the rows of the matrix 
mat_rows <- nrow(mat)
U <- mat
L <- diag(mat_rows)

#a is rows, b is columns
# Create loop to go through every row, column by column
for(a in 1:(nrow(mat) - 1)){
    for(b in ((a + 1):nrow(mat))){
    #Solve for L first, if not, the other values won't populate...
    L[a,b] <- (U[b,a]/U[a,a])
    U[b,] <- U[b,]-(U[a,] * (U[b,a]/U[a,a]))
    }
 }    
return(list("L" = L, "U" = U))
}


LU_Decomp(B)
LU_Decomp(A)
LU_Decomp(C)

```


