---
title: "607-Project-1"
author: "Gabe Abreu"
date: "2/20/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readr)
library(stringr)
library(ggplot2)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```

## Project 1

The purpose of this project is to text file and turn into a csv after its been "cleaned" and relevant data calculated and extracted. The exercise uses a chess tournament text with player names, state origin, win-loss record, pre-match ratings, post-match ratings.  

```{r tournament}
tournamentUrl <- "https://raw.githubusercontent.com/geeman1209/MSDATA2020/master/DATA607/Project1/tournamentinfo.txt"


rawData <- readLines(tournamentUrl)
View(rawData)
#Take a quick glimpse of the data
rawData[1:10]

#get rid of the dotted seperators 
lines <- str_detect(rawData, "-----------")
tournieData <- rawData[!lines]

#Eliminate Headers//Will Add Later
tournieData2 <- tournieData[-c(1:2)]
```
## Break down individual data components 

This is the most important and involved aspect of the project. My approach is to extract all the relevant components and piece it together into a final data frame. Since the text file dedicates two rows of data per player, you can extract information row by row and then further pull data from each row.


```{r}
#Get the rows with the names, id's, points, opponent id
id_name_points <- tournieData2[seq(from=1, to = length(tournieData2), by=2)]
head(id_name_points)

player_id <- c(1:64)

#Get rows with the states, pre-ratings, and post-ratings
st_ratings <- tournieData2[seq(from=2, to = length(tournieData2), by=2)]
head(st_ratings)


#Extract the data we need to do further calculations and create a new data frame

#solution to player_names found on stackoverflow
player_names <- str_extract(id_name_points, '[:alpha:](\\w+(?:\\s+\\w+)*)')
head(player_names)

player_points <- as.numeric(str_extract(id_name_points, "\\d.\\d"))
head(player_points)

player_state <- str_extract(st_ratings, '[:alpha:]{2}')
head(player_state)


#Detect and extract pre-ratings
both_ratings<-str_extract_all(st_ratings,"(( \\:)|(\\>))?.?\\d{1,}P*\\.?")

pre<-str_detect(unlist(both_ratings),"\\b\\d{3,4}P?\\b")
post<-str_detect(unlist(both_ratings),"\\>.?\\b\\d{3,4}P?\\b")

preRate<-unlist(both_ratings)[(pre)&(!post)]
preRate<-as.integer(str_replace_all(preRate,"P",""))

oppId <- str_extract_all(str_extract_all(id_name_points, "\\d+\\|"), "\\d+")
```

```{r}
#Calculate total number of opponents per player
num.opp <- sapply(oppId, length)
#Create a sequence up to the number of maximum opponents -> 7
max_seq_opp <- seq_len(max(num.opp))

#Create dataframe of opponent ids per match per player
#The idea is to replace the opponent ids with their rating and then take the average
mat <- t(sapply(oppId, "[", i = max_seq_opp)) 
DF_opp <- as.data.frame(mat)


key <- data.frame(player_id, preRate)
Copy_dfOpp <- DF_opp
Copy_dfOpp[] <- key$preRate[match(unlist(Copy_dfOpp), key$player_id)]
as_tibble(head(Copy_dfOpp))

#Make na.rm equal to true to handle NA values
Copy_dfOpp$Avg_Rating_Opp <- round(apply(Copy_dfOpp, 1, mean, na.rm=TRUE))
as_tibble(head(Copy_dfOpp))
```

```{r finalDF}
Final_DataFrame <- tibble(player_id, player_names, player_state, player_points, preRate, Copy_dfOpp$Avg_Rating_Opp)

colnames(Final_DataFrame) <- c("ID", "Names", "State", "Points", "Pre-Rating", "Average Opponent Pre-Rating")

as_tibble(Final_DataFrame)
```
## Write CSV File
```{r csv}
write.csv(Final_DataFrame, "clean_tournament_info.csv", row.names = FALSE)
```
## Graphical Analysis


```{r graphs, echo= TRUE}

ggplot(Final_DataFrame, aes(Points, `Average Opponent Pre-Rating`, color="player id")) + geom_point(color="green") + geom_text(aes(label=player_id),hjust=0, vjust=0) + theme(
legend.justification = c("right", "top"))

ggplot(Final_DataFrame, aes( `Average Opponent Pre-Rating`, preRate, color="players")) + geom_point(color="black") + theme(
legend.justification = c("right", "top"))
```

# Conclusion

This was a harder assignment than first anticipated. It took a lot of online searching to find the best way to approach all the text data and try/test the correct regex to detect and extract the relevant information. 


# Sources/References:

Solution to find names:
https://stackoverflow.com/questions/37307727/regex-to-extract-contact-name-and-email-from-string

A nice guide to cleaning text files:
https://cran.r-project.org/doc/contrib/de_Jonge+van_der_Loo-Introduction_to_data_cleaning_with_R.pdf

Stringr cheat sheet:
https://github.com/rstudio/cheatsheets/blob/master/strings.pdf

